\morus: Biases in the Stream

### 2018-05-10 morusAC_04_Mini_trails.tex
working on introduction
homogenized references
minor consistency fix: MORUS

\section{Linear Trail for \cipher{MiniMORUS}}
\label{sec/minitrails}

In this section, we describe how we build a trail for \cipher{MiniMORUS} before computing its bias and experimentally verify it.

@@

To build a linear trail on \cipher{MiniMORUS}, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
\begin{itemize}
    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
\end{itemize}

@@

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, alpha]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
      ;
    \draw[trail, alpha, dashed]
      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
      ;
  \end{tikzpicture}%}}}
  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, beta]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (tlll-1) -- (W-20) node[above] {$i$}
      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
      (xnd-1) -- (and-1)
      (W40) node[below] {\phantom{$i$}}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, gamma]
      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
      (xnd1) -- (and1)
      (xor1) -- (txor1)
      (txor1) -- (W-24) node[above] {$i$}
      (lll1) -- (W41) node[below] {$i+b_1$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
  \end{subfigure}
  \bigskip

  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, delta]
      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
      (lll4) -- (W44) node[below] {$i+b_4$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, epsil]
      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
      (lll2) -- (W42) node[below] {$i+b_2$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
    \small
    \begin{align*}
      \intertext{\cipher{MiniMORUS-640}:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
      \intertext{\cipher{MiniMORUS-1280}:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
    \end{align*}
    \caption*{\cipher{MiniMORUS} instances}
  \end{subfigure}
  \caption{\cipher{MiniMORUS} linear trail fragments.}
  \label{fig:minimorusfragments}
\end{figure}

@@

In order to build the equation for the full trail, we start with $E^2_0$:
\[
S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
\]
In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$). This yields:
\begin{align*}
&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
\end{align*}
We then need to cancel two terms of the form $S^t_{2,i}$. To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$. This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms. By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms. By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms). Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits. Ultimately, for \cipher{MiniMORUS1280}, this yields the equation:
\begin{flalign*}
&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
\end{flalign*}

The equation for \cipher{MiniMORUS640} is very similar, and is given in \Cref{sec:traileq}.

@@

In the equation for \cipher{MiniMORUS1280} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1. By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16. Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$. The same holds for \cipher{MiniMORUS640} (cf. \Cref{sec:traileq}).

@@

To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}. In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$). Results are displayed on \Cref{tab:miniapproximations}.
While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}. The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.

%\noindent
%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
%\begin{align*}
%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
%\intertext{Combined ciphertext-only approximation:}
%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
%\end{align*}
%\bigskip
%
%\noindent
%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
%\begin{align*}
%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
%\intertext{Combined ciphertext-only approximation:}
%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
%\end{align*}

\begin{table}[h!]
  \caption{Experimental verification of trail biases.}
  \label{tab:miniapproximations}
  \centerline{
  \begin{tabular}{@{}llSSS@{}}
    \toprule
    & & \multicolumn{3}{@{}c@{}}{Weight} \\
    \cmidrule{3-5}
    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS640}}
             & {Exp.} & {Bool.} & {Meas.} \\
    \midrule
    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
             & 7    & 7     & 7 \\
    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
             & 9    & 9     & 9 \\
    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
             & 16   & 16    & 15.5 \\
    \midrule
    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS1280}} \\
    \midrule
    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
             & 7    & 7     & 7 \\
    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
             & 9    & 9     & 9 \\
    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
             & 16   & 16    & 15.9 \\
    \bottomrule
  \end{tabular}
  }
\end{table}

>>>>

%!TEX root = morusAC.tex

\section{Linear Trail for \MiniMORUS}
\label{sec/minitrails}

In this section, we describe how we build a trail for \MiniMORUS before computing its bias and experimentally verify it.

@@

To build a linear trail on \MiniMORUS, we combine the following five trail fragments $\alpha^t_i$, $\beta^t_i$, $\gamma^t_i$, $\delta^t_i$, $\varepsilon^t_i$, where the subscript $i$ denotes a bit position, and the superscript $t$ denotes a step number:
\begin{itemize}
    \item $\alpha^t_i$ approximates (one bit of) state word $S_0$ using the ciphertext;
    \item $\beta^t_i$ approximates $S_1$ using $S_0$ and the ciphertext;
    \item $\gamma^t_i$ approximates $S_4$ using two approximations of $S_1$ in consecutive steps;
    \item $\delta^t_i$ approximates $S_2$ using two approximations of $S_4$ in consecutive steps;
    \item $\varepsilon^t_i$ approximates $S_0$ using two approximations of $S_2$ in consecutive steps.
\end{itemize}

@@

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{subfigure}{.32\textwidth}
  \centering
  \tikzsetnextfilename{minimorus_part_alpha}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, alpha]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
      ;
    \draw[trail, alpha, dashed]
      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
      ;
  \end{tikzpicture}%}}}
  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \centering
  \tikzsetnextfilename{minimorus_part_beta}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, beta]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (tlll-1) -- (W-20) node[above] {$i$}
      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
      (xnd-1) -- (and-1)
      (W40) node[below] {\phantom{$i$}}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \tikzsetnextfilename{minimorus_part_gamma}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, gamma]
      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
      (xnd1) -- (and1)
      (xor1) -- (txor1)
      (txor1) -- (W-24) node[above] {$i$}
      (lll1) -- (W41) node[below] {$i+b_1$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
  \end{subfigure}
  \bigskip

  \begin{subfigure}{.32\textwidth}
  \tikzsetnextfilename{minimorus_part_delta}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, delta]
      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
      (lll4) -- (W44) node[below] {$i+b_4$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \tikzsetnextfilename{minimorus_part_epsil}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, epsil]
      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
      (lll2) -- (W42) node[below] {$i+b_2$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
    \small
    \begin{align*}
      \intertext{\MiniMORUS[640]:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
      \intertext{\MiniMORUS[1280]:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
    \end{align*}
    \caption*{\MiniMORUS instances}
  \end{subfigure}
  \caption{\MiniMORUS linear trail fragments.}
  \label{fig:minimorusfragments}
\end{figure}

@@

In order to build the equation for the full trail, we start with $E^2_0$:
\[
S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
\]
In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$). This yields:
\begin{align*}
&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
\end{align*}
We then need to cancel two terms of the form $S^t_{2,i}$. To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$. This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms. By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms. By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms). Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits. Ultimately, for \MiniMORUS[1280], this yields the equation:
\begin{flalign*}
&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
\end{flalign*}

The equation for \MiniMORUS[640] is very similar, and is given in \Cref{sec:traileq}.

@@

In the equation for \MiniMORUS[1280] from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1. By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16. Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$. The same holds for \MiniMORUS[640] (cf. \Cref{sec:traileq}).

@@

To confirm that our analysis is correct, we ran experiments on an implementation of \MiniMORUS[1280] and \MiniMORUS[640]. In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$). Results are displayed on \Cref{tab:miniapproximations}.
While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \MORUS[640]. The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.

%\noindent
%2 approximations for $S^{2,2}_{0}$ of \MiniMORUS[640]:
%\begin{align*}
%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
%\intertext{Combined ciphertext-only approximation:}
%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
%\end{align*}
%\bigskip
%
%\noindent
%2 approximations for $S^{2,2}_{0}$ of \MiniMORUS[1280]:
%\begin{align*}
%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
%\intertext{Combined ciphertext-only approximation:}
%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
%\end{align*}

\begin{table}[h!]
  \caption{Experimental verification of trail biases.}
  \label{tab:miniapproximations}
  \centerline{
  \begin{tabular}{@{}llSSS@{}}
    \toprule
    & & \multicolumn{3}{@{}c@{}}{Weight} \\
    \cmidrule{3-5}
    \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[640]}
             & {Exp.} & {Bool.} & {Meas.} \\
    \midrule
    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
             & 7    & 7     & 7 \\
    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
             & 9    & 9     & 9 \\
    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
             & 16   & 16    & 15.5 \\
    \midrule
    \multicolumn{2}{@{}l}{Approximations for \MiniMORUS[1280]} \\
    \midrule
    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
             & 7    & 7     & 7 \\
    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
             & 9    & 9     & 9 \\
    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
             & 16   & 16    & 15.9 \\
    \bottomrule
  \end{tabular}
  }
\end{table}
